---
title: "Big Data & AI code file"
output:
  pdf_document: default
  html_document: default
date: "2024-04-12"
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```
```{r, include = FALSE}
library(tidyverse)
library(tidytable)
library(caret)
library(stringr)
library(xgboost)
library(pROC)
library(tpc)
library(fixest)



df <- read.csv("/cloud/project/data/airbnb-project-msba-kaggle-train.csv")
```

```{r, echo=FALSE}
dt <- df

# Remove dollar signs
remove_dollar_sign <- c("price", "weekly_price", "monthly_price", "extra_people")

dt <-
  dt %>%
  mutate_at(remove_dollar_sign, ~str_replace_all(., pattern="\\$", replacement="")) %>%
  mutate_at(remove_dollar_sign, ~str_replace_all(., pattern=",", replacement="")) %>%
  mutate_at(remove_dollar_sign, ~as.numeric(.))

# Remove percentage signs
remove_pct_sign <- c("host_response_rate", "host_acceptance_rate")

dt <-
  dt %>%
  mutate_at(remove_pct_sign, ~str_replace_all(., pattern="\\%", replacement="")) %>%
  mutate_at(remove_pct_sign, ~as.numeric(.)) %>%
  mutate_at(remove_pct_sign, ~./100)

# Clean Up host_response_time
dt <- dt %>% 
  mutate(host_response_time = ifelse(host_response_time == "a few days or more" , 0, ifelse(host_response_time == "within a day", 1, ifelse(host_response_time == "within a few hours", 2, ifelse(host_response_time == "within an hour", 3, NA)))))

# quantifying the amenities column
dt$amenities = str_count(dt$amenities, ",") + 1

# merging bathrooms and bathrooms_text
dt$bathrooms_number = as.numeric(gsub("[^0-9\\.]", "", dt$bathrooms_text))
dt$bathrooms_merged <- coalesce(dt$bathrooms, dt$bathrooms_number)
dt <- dt %>% select(!c(bathrooms,bathrooms_text,bathrooms_number))

#standarardize weekly and monthly prices
dt <- dt%>% 
  mutate(weekly_price = ifelse(is.na(weekly_price), price * 7, weekly_price),
         monthly_price = ifelse(is.na(monthly_price), weekly_price * 4, monthly_price))

#Clean up room_type

# Perform one-hot encoding
encoded_data <- model.matrix(~ room_type - 1, data = dt)

# Convert the encoded data to a data frame
encoded_df <- as.data.frame(encoded_data)

# Rename the columns to remove the intercept term
colnames(encoded_df) <- gsub("^room_type", "", colnames(encoded_df))

# Add the one-hot encoded columns to the original data frame
dt <- cbind(dt, encoded_df) 

dt <- dt %>% select(!room_type)

# Clean bed_type column
dt <- dt %>% mutate(bed_type = ifelse(is.na(bed_type), "Not-Listed", bed_type))

# Perform one-hot encoding
encoded_data <- model.matrix(~ bed_type - 1, data = dt)

# Convert the encoded data to a data frame
encoded_df <- as.data.frame(encoded_data)

# Rename the columns to remove the intercept term
colnames(encoded_df) <- gsub("^bed_type", "", colnames(encoded_df))

# Add the one-hot encoded columns to the original data frame
dt <- cbind(dt, encoded_df)

dt1 <- dt %>% 
  select(!c(`Not-Listed`, bed_type))
```


### Overall Market Model
```{r, echo = FALSE}
dt <- dt1 %>%
  select(-longitude, -latitude,-region_name, - region_parent_name, -host_listing_count, -square_feet, -"Entire home/apt", -"Hotel room", -"Private room", -"Shared room", -"Pull-out Sofa", -"Real Bed" )

# Calculate the mean price from non-missing values
mean_price <- mean(dt$price, na.rm = TRUE)

# Impute missing values in the 'price' column with this mean
dt <- dt %>%
  mutate(price = ifelse(is.na(price), mean_price, price))

# Impute missing values in 'weekly_price' and 'monthly_price'
dt <- dt %>%
  mutate(weekly_price = ifelse(is.na(weekly_price), price * 7, weekly_price),
         monthly_price = ifelse(is.na(monthly_price), weekly_price * 4, monthly_price))

# Impute missing values in all numeric columns with the mean
dt <- dt %>%
  mutate(across(where(is.numeric), ~ ifelse(is.na(.), mean(., na.rm = TRUE), .)))

dfc<-dt

# Remove any variables that are string (character)
dt <-
  dt %>%
  select_if(~!is.character(.))

# This selects the 'host_response_time' and all other columns that are not characters.
dt <- dt %>% select(host_response_time, where(~ !is.character(.))) 

set.seed(3.14159)

index <- createDataPartition(dt$high_booking, p = 0.8, list = FALSE)
train.st<- dt[index,]
test.st <- dt[-index,]

data_train=as.matrix(train.st%>% select(!high_booking)) 
data_test=as.matrix(test.st%>% select(!high_booking)) 

data_2train <- xgb.DMatrix(data_train, label = train.st$high_booking)
data_2test <- xgb.DMatrix(data_test, label = test.st$high_booking)

Boost1 <- xgb.train(data = data_2train, max_depth = 3,
              eta = 1, nthread = 5, nrounds = 6,
              watchlist = list(train = data_2train, eval = data_2test),
              objective = "binary:logistic")

pred=ifelse(predict(object = Boost1,data_2test)>0.5,"1","0")

roc(test.st$high_booking, predict(object = Boost1,data_2test))$auc
```

```{r, echo = FALSE}
importance_matrix <- xgb.importance(feature_names = colnames(data_train), model = Boost1)

# Plot feature importance
xgb.plot.importance(importance_matrix)
```

### LA Market Model
```{r, echo = FALSE}
dt <- dt1 %>% 
  filter(market == "Los Angeles")

dt <- dt %>%
  select(-longitude, -latitude,-region_name, - region_parent_name, -host_listing_count, -square_feet, -"Entire home/apt", -"Hotel room", -"Private room", -"Shared room", -"Pull-out Sofa", -"Real Bed" )

# Calculate the mean price from non-missing values
mean_price <- mean(dt$price, na.rm = TRUE)

# Impute missing values in the 'price' column with this mean
dt <- dt %>%
  mutate(price = ifelse(is.na(price), mean_price, price))

# Impute missing values in 'weekly_price' and 'monthly_price'
dt <- dt %>%
  mutate(weekly_price = ifelse(is.na(weekly_price), price * 7, weekly_price),
         monthly_price = ifelse(is.na(monthly_price), weekly_price * 4, monthly_price))

# Impute missing values in all numeric columns with the mean
dt <- dt %>%
  mutate(across(where(is.numeric), ~ ifelse(is.na(.), mean(., na.rm = TRUE), .)))

# Remove any variables that are string (character)
dt <-
  dt %>%
  select_if(~!is.character(.))

# This selects the 'host_response_time' and all other columns that are not characters.
dt <- dt %>% select(host_response_time, where(~ !is.character(.))) 


set.seed(3.14159)

index <- createDataPartition(dt$high_booking, p = 0.8, list = FALSE)
train.st<- dt[index,]
test.st <- dt[-index,]

data_train=as.matrix(train.st%>% select(!high_booking)) 
data_test=as.matrix(test.st%>% select(!high_booking)) 

data_2train <- xgb.DMatrix(data_train, label = train.st$high_booking)
data_2test <- xgb.DMatrix(data_test, label = test.st$high_booking)

Boost1 <- xgb.train(data = data_2train, max_depth = 3,
              eta = 1, nthread = 5, nrounds = 6,
              watchlist = list(train = data_2train, eval = data_2test),
              objective = "binary:logistic")

pred=ifelse(predict(object = Boost1,data_2test)>0.5,"1","0")

library(pROC)

roc(test.st$high_booking, predict(object = Boost1,data_2test))$auc
```

```{r, echo = FALSE}
importance_matrix <- xgb.importance(feature_names = colnames(data_train), model = Boost1)

# Plot feature importance
xgb.plot.importance(importance_matrix)
```
### Causal Inference
```{r, echo = FALSE}
selected_columns <- c("extra_people", "host_response_rate", "accommodates", "price","requires_license","reviews_per_month", "high_booking")

dfc_dag <- dfc %>%
  filter(market == "Los Angeles") %>% 
  select(selected_columns)

selected_columns <- c("extra_people", "host_response_rate", "accommodates", "price","requires_license","reviews_per_month", "high_booking", "zipcode")

dfc <- dfc %>%
  filter(market == "Los Angeles") %>% 
  select(selected_columns)
```

```{r, echo = FALSE}
dfc_dag$high_booking <- as.numeric(dfc_dag$high_booking)
suffStat <- list(C = cor(dfc_dag), n = nrow(dfc_dag))

varNames <- colnames(dfc_dag)

exogenous_vars <- c("host_response_rate", "price","reviews_per_month")

# Run the tpc function with the selected exogenous variables
skel.dfc <- tpc(suffStat, indepTest = gaussCItest, labels = varNames, alpha = 0.01, context.all = exogenous_vars)


#str(dt)
```

```{r, echo = FALSE}
 plot(skel.dfc, main = "TPC Estimate with Global ContextVariables", cex = 100.5)
#names(dfc)
```
```{r, echo = FALSE}
library(corrplot)
cor_dfc <- dfc %>% 
  select(!c(zipcode, requires_license))

corrplot(cor(cor_dfc, use = "pairwise.complete.obs"), type = "upper")
```


```{r, echo = FALSE}
# feols(high_booking ~ accommodates | price, data = dfc)
```
```{r, echo = FALSE}
acc_quartile_breaks <- quantile(dfc$accommodates, probs = seq(0, 1, by = 0.25), na.rm = TRUE)
price_quaritle_breaks <- quantile(dfc$price, probs = seq(0, 1, by = 0.25), na.rm = TRUE)
hrr_quaritle_breaks <- quantile(dfc$host_response_rate, probs = seq(0, 1, by = 0.25), na.rm = TRUE)

# Create the numeric categories based on quartile breaks
dfc$acc_range <- cut(dfc$accommodates, breaks = acc_quartile_breaks, include.lowest = TRUE, labels = FALSE)
dfc$price_range <- cut(dfc$price, breaks = price_quaritle_breaks, include.lowest = TRUE, labels = FALSE)
```
```{r, echo = FALSE}
# dfc1 <- dfc %>%
#   mutate(accommodates = ifelse(acc_range != 1, 0, accommodates))
```
```{r, echo = FALSE}
# feols(high_booking ~ accommodates |price, data = dfc1)

```
```{r, echo = FALSE}
# dfc2 <- dfc %>%
  # mutate(accommodates = ifelse(acc_range != 2, 0, accommodates))
```

```{r, echo = FALSE}
# feols(high_booking ~ accommodates | price, data = dfc2)
```
```{r, echo = FALSE}
# dfc3 <- dfc %>%
#   mutate(accommodates = ifelse(acc_range != 3, 0, accommodates))
# feols(high_booking ~ accommodates | price, data = dfc3)
```
```{r, echo = FALSE}
# dfc4 <- dfc %>%
#   mutate(accommodates = ifelse(acc_range != 4, 0, accommodates))
# feols(high_booking ~ accommodates | price, data = dfc4)
```

```{r, echo = FALSE}
# x <- dfc %>%
#   filter(acc_range==2)
# x$price_range <- NA
# quartile_breaks <- quantile(x$price, probs = seq(0, 1, by = 0.25), na.rm = TRUE)
# 
# # Create the numeric categories based on quartile breaks
# x$price_range <- cut(x$price, breaks = quartile_breaks, include.lowest = TRUE, labels = FALSE)
# 
# hist(x$price,breaks=500,xlim=range(0,300))
```

```{r, echo = FALSE}
# Create the numeric categories based on quartile breaks
# quartile_breaks <- quantile(x$extra_people, probs = seq(0, 1, by = 0.50), na.rm = TRUE)
# x$ep_range <- cut(x$extra_people, breaks = quartile_breaks, include.lowest = TRUE, labels = FALSE)
# head(x)
```

```{r, echo = FALSE}
# x4 <- x %>%
#   mutate(extra_people = ifelse(ep_range != 1, 0, extra_people))
# feols(high_booking ~ extra_people | price + host_response_rate + reviews_per_month, data = x4)
```
```{r, echo = FALSE}
# x5 <- x %>%
  # mutate(extra_people = ifelse(ep_range != 2, 0, extra_people))
# feols(high_booking ~ extra_people | price + host_response_rate + reviews_per_month, data = x5)
```


```{r, echo = FALSE}
# host_df <- dfc %>% 
#   filter(acc_range ==2) %>% 
#   select(high_booking, host_response_rate, reviews_per_month, price) %>% 
#   na.omit() %>% 
#   mutate(th = ifelse(host_response_rate <= quantile(host_response_rate, .05, na.rm = T), 1, 0)) 
# 
# feols(high_booking~th| reviews_per_month + price, data = host_df)
```

```{r, echo = FALSE}
# hist(host_df$host_response_rate, breaks = 20)
# 
# quantile(host_df$host_response_rate, seq(0, .25, by = .05))
```
### Accommodates Findings
```{r, echo = FALSE}

####### Origonal Version ###############


# host_df <- dfc %>% 
#   filter(acc_range ==2) %>% 
#   select(high_booking, host_response_rate, reviews_per_month, price) %>% 
#   na.omit() %>% 
#   mutate(th = ifelse(host_response_rate >= quantile(host_response_rate, .05, na.rm = T), 1, 0)) 
# 
# feols(high_booking~th| reviews_per_month + price, data = host_df)


########### Revised Version ###################

# Accommodates Range
hist(dfc$accommodates)
```
```{r, echo = FALSE}
feols(high_booking~accommodates|zipcode , data = dfc)
```
Significant and negative. Recommend investing in airbnbs that accomodates less people. 

```{r, echo = FALSE}
dfc %>%
  group_by(high_booking, acc_range) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  group_by(acc_range) %>%
  mutate(prop = n / sum(n)) %>%
  ggplot(aes(x = factor(acc_range), y = n, fill = factor(high_booking))) +
  geom_col(position = "dodge") +
  geom_text(aes(label = round(prop, 2)), 
            position = position_dodge(width = 0.9), 
            vjust = -0.25) +  # Adjust vjust for label positioning
  ggtitle("Count and Proportion by Accomodates Range and High Booking") +
  xlab("Accomodates Range") +
  ylab("Count") +
  scale_fill_discrete(name = "High Booking")
```
```{r, echo = FALSE}
quantile(dfc$accommodates)
```
From the chart and these quantiles, we can see that we want to invest in Q1 (accommodates for 1-2 people) and Q2 (accommodates for 4-5 people)
### Price Findings
```{r, echo = FALSE}
price1_df <- dfc %>% 
  filter(acc_range == 1)

hist(price1_df$price, breaks = 1000, xlim = c(0, 250))
```
```{r, echo = FALSE}
# price1_mod <- glm(high_booking~price, data = price1_df, family = "binomial")
# 
# summary(price1_mod)

feols(high_booking~price|zipcode,data = price1_df)
```
Signifiant and negative. When investing in ABNB that accommodates for 1-2 people, try not to charge too much. 

```{r, echo = FALSE}
dfc %>%
  filter(acc_range == 1) %>%
  group_by(high_booking, price_range) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  group_by(price_range) %>%
  mutate(prop = n / sum(n)) %>%
  ggplot(aes(x = factor(price_range), y = n, fill = factor(high_booking))) +
  geom_col(position = "dodge") +
  geom_text(aes(label = round(prop, 2)), 
            position = position_dodge(width = 0.9), 
            vjust = -0.25) +  # Adjust vjust for label positioning
  ggtitle("Count and Proportion by Price Range and High Booking") +
  xlab("Price Range") +
  ylab("Count") +
  scale_fill_discrete(name = "High Booking")
```
```{r, echo = FALSE}
quantile(price1_df$price)
```
From the graph and the quantiles make sure not to charge over 110 for ABNBs that accommodates 1-2 people
```{r, echo = FALSE}
price3_df <- dfc %>% 
  filter(acc_range == 3)

hist(price3_df$price, breaks = 100)
```


```{r, echo = FALSE}
# price3_mod <- glm(high_booking~price, data = price3_df, family = "binomial")
# 
# summary(price3_mod)


feols(high_booking~price|zipcode,data = price3_df)
```
Not significant
```{r, echo = FALSE}
dfc %>%
  filter(acc_range == 3) %>%
  group_by(high_booking, price_range) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  group_by(price_range) %>%
  mutate(prop = n / sum(n)) %>%
  ggplot(aes(x = factor(price_range), y = n, fill = factor(high_booking))) +
  geom_col(position = "dodge") +
  geom_text(aes(label = round(prop, 2)), 
            position = position_dodge(width = 0.9), 
            vjust = -0.25) +  # Adjust vjust for label positioning
  ggtitle("Count and Proportion by Price Range and High Booking") +
  xlab("Price Range") +
  ylab("Count") +
  scale_fill_discrete(name = "High Booking")

```
```{r, echo = FALSE}
quantile(price3_df$price)
```

No specific recommendation for this level of accommodates. Price proportion is similar across quantiles however Q1 has the lowest prop. 

### Host Response Rate Findings

```{r, echo = FALSE}
host1_df <- dfc %>% 
  filter(acc_range == 1)
```

```{r, echo = FALSE}
# host1_mod <- glm(high_booking~host_response_rate, data = host1_df, family = "binomial")
# 
# summary(host1_mod)

feols(high_booking~host_response_rate|zipcode,data = host1_df)
```
Positive, however not very significant.
```{r, echo = FALSE}
dfc %>%
  filter(acc_range == 1) %>%
  mutate(hrr_range = ifelse(host_response_rate > .75, 1, 0)) %>% 
  group_by(high_booking, hrr_range) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  group_by(hrr_range) %>%
  mutate(prop = n / sum(n)) %>%
  ggplot(aes(x = factor(hrr_range), y = n, fill = factor(high_booking))) +
  geom_col(position = "dodge") +
  geom_text(aes(label = round(prop, 2)), 
            position = position_dodge(width = 0.9), 
            vjust = -0.25) +  # Adjust vjust for label positioning
  ggtitle("Count and Proportion by Price Range and High Booking") +
  xlab("Price Range") +
  ylab("Count") +
  scale_fill_discrete(name = "High Booking")
```
Above 75% response rate improves high booking for ABNBs that accommodates for 1-2 people. 

```{r, echo = FALSE}
quantile(host1_df$host_response_rate)
```

```{r, echo = FALSE}
host3_df <- dfc %>% 
  filter(acc_range == 3)
```

```{r, echo = FALSE}
feols(high_booking~host_response_rate|zipcode,data = host3_df)
```
Not sig. 
```{r, echo = FALSE}
dfc %>%
  filter(acc_range == 3) %>%
  mutate(hrr_range = ifelse(host_response_rate > .75, 1, 0)) %>% 
  group_by(high_booking, hrr_range) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  group_by(hrr_range) %>%
  mutate(prop = n / sum(n)) %>%
  ggplot(aes(x = factor(hrr_range), y = n, fill = factor(high_booking))) +
  geom_col(position = "dodge") +
  geom_text(aes(label = round(prop, 2)), 
            position = position_dodge(width = 0.9), 
            vjust = -0.25) +  # Adjust vjust for label positioning
  ggtitle("Count and Proportion by Price Range and High Booking") +
  xlab("Price Range") +
  ylab("Count") +
  scale_fill_discrete(name = "High Booking")
```
Above 75% response rate improves high booking for ABNBs that accommodates 4-5 people. 
